---
title: "MicrobiomeR: Analysis"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Data Wrangling

```{r message=FALSE, warning=FALSE}
# Get the data files from package
input_files <- pkg.data$input_files
biom_file <- input_files$biom_files$silva # Path to silva biom file
tree_file <- input_files$tree_files$silva # Path to silva tree file
metadata_file <- input_files$metadata # Path to Nephele metadata
parse_func <- parse_taxonomy_silva_128 # A custom phyloseq parsing function for silva annotations

# Get the phyloseq object
phy_obj <- get_phyloseq_obj(biom_file = biom_file, 
                                    tree_file = tree_file, 
                                    metadata_file = metadata_file,
                                    parse_func = parse_func)

# Get the taxmap object in the raw format
raw_metacoder <- as_MicrobiomeR_format(obj = phy_obj, format = "raw_format")
```

# Filtering

```{r message=FALSE, warning=FALSE}
# Remove Archaea from the metacoder object
metacoder_obj <- taxa::filter_taxa(obj = raw_metacoder,
                                   taxon_names == "Archaea",
                                   subtaxa = TRUE,
                                   invert = TRUE)

# Ambiguous Annotation Filter - Remove taxonomies with ambiguous names
metacoder_obj <- metacoder::filter_ambiguous_taxa(metacoder_obj, 
                                                  subtaxa = TRUE)

# Low Sample Filter - Remove the low samples
# The sample filter should generally be implemented first
metacoder_obj <- sample_filter(obj = metacoder_obj,
                               .f_filter = ~sum(.),
                               .f_condition = ~.>= 20, 
                               validated = TRUE)

# Master Threshold Filter - Add the otu_proportions table and then filter OTUs based on min %
metacoder_obj <- otu_proportion_filter(obj = metacoder_obj,
                                       otu_percentage = 0.00001)

# Taxon Prevalence Filter - Add taxa_abundance and taxa_proportions and then filter OTUs that do not
# appear more than a certian amount of times in a certain percentage of samples at the specified
# agglomerated rank.  This is considered a supervised method, because it relies on intermediate
# taxonomies to filter the data.
# The default minimum abundance is 5 and the sample percentage is 0.5 (5%).
# Phylum
metacoder_obj <- taxa_prevalence_filter(obj = metacoder_obj,
                                        rank = "Phylum")
# Class
metacoder_obj <- taxa_prevalence_filter(obj = metacoder_obj,
                                        rank = "Class", 
                                        validated = TRUE)
# Order
metacoder_obj <- taxa_prevalence_filter(obj = metacoder_obj,
                                        rank = "Order", 
                                        validated = TRUE)

# OTU Prevalence Filter - Filter OTUs that do not appear more than a certian amount of times in a
# certain percentage of samples.  This is considered an unsupervised method, because it relies only
# on the leaf OTU ids to filter the data.
metacoder_obj <- otu_prevalence_filter(obj = metacoder_obj, 
                                       validated = TRUE)

# Coefficient of Variation Filter - Filter OTUs based on the coefficient of variation
metacoder_obj <- cov_filter(obj = metacoder_obj,
                            coefficient_of_variation = 3, 
                            validated = TRUE)
```

# Analysis

Analysis is primarily done with metacoder, MicrobiomeR, and ggplot2.  Before beginning the analysis
it's wise to create an output directory.  Use `end_path=FALSE` to avoid the creation of a date 
formatted directory.

```{r message=FALSE, warning=FALSE}
# Create a directory for whichever plot you want to save
htp <- get_output_dir(end_path = FALSE, start_path = "output", plot_type = "heat_tree")
cpp <- get_output_dir(end_path = FALSE, start_path = "output", plot_type = "correlation")
```

## Statistics

Statistical analysis is primarily done with the help of metacoder style functions such as the 
`calc_*()` group of functions, and `compare_groups()`.  The taxa function `taxonomy_table()` is also 
useful for matching stats with the proper taxonomic annotation.  MicrobiomeR creates the proper 
tables with `as_MicrobiomeR_format(format = "analyzed_format", ...)`.

```{r message=FALSE, warning=FALSE}
# Get the statistical observation data.
metacoder_obj <- as_MicrobiomeR_format(obj = metacoder_obj, format = "analyzed_format")
```

