---
title: "Using Color Palettes"
author: "Rob Gilmore & Shaurita Hutchins"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Color Palettes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(MicrobiomeR)
pkg.data <- MicrobiomeR:::pkg.private
```

# Introduction

The color palette workflow created for MicrobiomeR are based on the 
`get_color_palette(pal_func = ...)` with a palette function that returns
`grDevices::colorRampPalette(...)` (called), `grDevices::colorRampPalette` (not called),
or a color palette function that returns a character vector.  Several palette functions
come built with MicrobiomeR and are used in the plotting functions.

## Getting a Color Palette

Getting a color palette is simple.  You can toggle the palette preview with `display=FALSE`.

```{r message=FALSE, warning=FALSE}
analyzed_silva <- as_MicrobiomeR_format(MicrobiomeR::raw_silva_2, "analyzed_format")
data <- analyzed_silva$data$stats_tax_data$Phylum
data_len <- length(unique(data))
custom_pal <- get_color_palette(color_no=data_len)
print(custom_pal)

```

## Scico

We have built in support for the [scico](https://github.com/thomasp85/scico) package.

```{r message=FALSE, warning=FALSE}
basic_pal <- scico_palette(scico_palette="hawaii")(25)
pie(rep(1, length(basic_pal)), col=basic_pal)
print(basic_pal)
better_pal <- get_color_palette(pal_func = scico_palette, color_no = 25)
print(better_pal)
```

## Viridis

We also provide support for the [viridis](https://github.com/sjmgarnier/viridis) package.

Typically, you'd use the viridis palette like we do below.
```{r message=FALSE, warning=FALSE}
basic_pal <- viridis::viridis(n=25, option="magma")
pie(rep(1, length(basic_pal)), col=basic_pal)
print(basic_pal)
```

With `MicrobiomeR`, the viridis palette can be optimized.
```{r message=FALSE, warning=FALSE}
better_pal <- get_color_palette(pal_func = viridis_palette, color_no = 25)
pie(rep(1, length(better_pal)), col=better_pal)
print(better_pal)
```

## Combining Color Palettes

Here is a useful functionality.  Combining color palettes can help with situations
where you need something more custom.  This can be especially helpful if you are using
more than 20 colors.  The `combination_palette` function can take a dynamic number of 
arguments, must be named and the value must be a `list`.  The list members can contain members with the following names:  

> You can use any name for your arguments, but the values must be a named list.
  palette:  This is a palette function that returns a vector of colors.
  args:  This is another named list used for the palette function parameters.
  range:  This is a range (1:10) used to subset the color palette vector.
  rev:  This is a logical (TRUE/FALSE)used to reverse the color palette.

```{r message=FALSE, warning=FALSE}
# This is the code for MicrobiomeR::viridis_magma_palette, but
# it's also a useful example for understanding how this works.
 viridis_magma_palette <- function(viridis_number = 800, 
                                   viridis_range = 300:viridis_number,
                                   viridis_rev = TRUE,
                                   magma_number = 500, 
                                   magma_range = 0:magma_number, 
                                   magma_rev = FALSE, 
                                   ...) {
   # Get the arguments for both functions
   if (!missing(...)){
       v_args = list(n=viridis_number, ...)
       m_args = list(n=magma_number, ...)
   } else {
       v_args = list(n=viridis_number)
       m_args = list(n=magma_number)
   }
   # Get combo palette
   crp <- combination_palette(viridis =
                                 list(palette = viridis::viridis,
                                       args = v_args,
                                       range = viridis_range,
                                       rev = viridis_rev),
                                magma =
                                  list(palette = viridis::magma,
                                       args = m_args,
                                       range = magma_range,
                                       rev = magma_rev)
                                )
     return(crp)
 }
combo_palette <- get_color_palette(pal_func = viridis_magma_palette)
combo_palette
```
