---
title: "MicrobiomeR: Filtering"
author: "Robert Gilmore"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Filtering}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(MicrobiomeR)
pkg.data <- MicrobiomeR:::pkg.private
```

# Data Wrangling

```{r message=FALSE, warning=FALSE}
# Get the data files from package
input_files <- pkg.data$input_files
biom_file <- input_files$biom_files$silva # Path to silva biom file
tree_file <- input_files$tree_files$silva # Path to silva tree file
metadata_file <- input_files$metadata # Path to Nephele metadata
parse_func <- parse_taxonomy_silva_128 # A custom phyloseq parsing function for silva annotations

# Get the phyloseq object
phy_obj <- get_phyloseq_obj(biom_file = biom_file, 
                                    tree_file = tree_file, 
                                    metadata_file = metadata_file,
                                    parse_func = parse_func)

# Get the taxmap object in the raw format
raw_metacoder <- as_MicrobiomeR_format(obj = phy_obj, format = "raw_format")
```

# Taxmap Filtering

Metacoder is a highly useful package that comes with tons of features right out of the box.  Functions
such as `metacoder::filter_ambiguous_taxa()`, `taxa::filter_taxa()`, and `taxa::filter_obs()` for instance
can almost always be used in your workflow.  

```{r message=FALSE, warning=FALSE}
# Remove Archaea from the metacoder object
metacoder_obj <- taxa::filter_taxa(obj = raw_metacoder,
                                   taxon_names == "Archaea",
                                   subtaxa = TRUE,
                                   invert = TRUE)

# Ambiguous Annotation Filter - Remove taxonomies with ambiguous names
metacoder_obj <- metacoder::filter_ambiguous_taxa(metacoder_obj, 
                                                  subtaxa = TRUE)
```

# Basic Filtering

Three functions are provided by MicrobiomeR that do some basic filtering:

* `sample_id_filter()` for filtering samples.
* `taxon_id_filter()` for filtering by taxon_id, which includes intermediate taxa.
* `otu_id_filter()` for filtering by otu_id, which only includes leaf taxa.

These functions all take the same parameters, most noteably a transformation function (.f_transform),
a filtering function (.f_filter), and a coditional function (.f_condition).

```{r message=FALSE, warning=FALSE}
# Low Sample Filter - Remove the low samples
# The sample filter should generally be implemented first
metacoder_obj <- sample_filter(obj = metacoder_obj,
                               .f_filter = ~sum(.),
                               .f_condition = ~.>= 20, 
                               validated = TRUE)
```

# Advanced Filtering

The advanced filtering functions available with MicrobiomeR do several things.  They wrap the basic 
filtering functions mentioned above, they wrap common metacoder and taxa functions, and they mimic 
the tools found in the phyloseq package.  Below I've mentioned some of them and how they relate to 
the phyloseq package:

* `agglomerate_metacoder()` is equivalent to `phyloseq::tax_glom()`.
* `otu_proportion_filter()` is seen in the first step in phyloseq's [preprocessing](http://joey711.github.io/phyloseq/preprocess.html#preprocessing) vignette.
* `Prevalence Filtering`  functions filter observations by thier prevelance across samples.
    * (Unsupervised) `otu_prevelance_filter()` is seen in phyloseq's [prevalence
    filtering](http://web.stanford.edu/class/bios221/MicrobiomeWorkflowII.html#prevalence_filtering) vignette.
    * (Supervised) `taxa_prevalence_filter()` is seen in phyloseq's [taxonomic filtering](http://web.stanford.edu/class/bios221/MicrobiomeWorkflowII.html#taxonomic_filtering) vignette.
* `cov_filter()` is seen in the nineth step in phyloseq's [preprocessing](http://joey711.github.io/phyloseq/preprocess.html#preprocessing) vignette.

```{r message=FALSE, warning=FALSE}
# Master Threshold Filter - Add the otu_proportions table and then filter OTUs based on min %
metacoder_obj <- otu_proportion_filter(obj = metacoder_obj,
                                       otu_percentage = 0.00001)

# Taxon Prevalence Filter - Add taxa_abundance and taxa_proportions and then filter OTUs that do not
# appear more than a certian amount of times in a certain percentage of samples at the specified
# agglomerated rank.  This is considered a supervised method, because it relies on intermediate
# taxonomies to filter the data.
# The default minimum abundance is 5 and the sample percentage is 0.5 (5%).
# Phylum
metacoder_obj <- taxa_prevalence_filter(obj = metacoder_obj,
                                        rank = "Phylum")
# Class
metacoder_obj <- taxa_prevalence_filter(obj = metacoder_obj,
                                        rank = "Class", 
                                        validated = TRUE)
# Order
metacoder_obj <- taxa_prevalence_filter(obj = metacoder_obj,
                                        rank = "Order", 
                                        validated = TRUE)

# OTU Prevalence Filter - Filter OTUs that do not appear more than a certian amount of times in a
# certain percentage of samples.  This is considered an unsupervised method, because it relies only
# on the leaf OTU ids to filter the data.
metacoder_obj <- otu_prevalence_filter(obj = metacoder_obj, 
                                       validated = TRUE)

# Coefficient of Variation Filter - Filter OTUs based on the coefficient of variation
metacoder_obj <- cov_filter(obj = metacoder_obj,
                            coefficient_of_variation = 3, 
                            validated = TRUE)
```

# Other Filtering

As mentioned previously taxmap filtering can be done in any way that fits your needs with the `taxa`
and `metacoder` packages.  However, MicrobiomeR also provides some utility based function for 
filtering/manipulating your _observation_ data by hand.  Observation data can be accessed within
the metacoder object.  Make sure you don't manipulate existing data inside of your taxmap object
unless you're absolutely sure you know what you're doing.

## Transposing

```{r message=FALSE, warning=FALSE}
# Get the abundance tables from the taxmap object
taxa_abund <- metacoder_obj$data$taxa_abundance
otu_abund <- metacoder_obj$data$otu_abundance

# Transposing with one ID (taxon_id)
taxa_abund %>% transposer(ids = "taxon_id", header_name = "samples")

# Re-Transposing with one ID (taxon_id)
taxa_abund %>% transposer("taxon_id", "samples") %>% transposer("samples", "taxon_id")

# Transposing with two IDs (taxon_id, otu_id)
# The column headers will be a combination of the categorical data that will
# be parsed and split back into individual columns if retransposed.
otu_abund %>% transposer("otu_id", "samples")

# Re-Transposing with two IDs (taxon_id, otu_id)
# When you transpose categorical data, the column headers are lost.
# To re-transpose you have to supply these headers
otu_abund %>% transposer("otu_id", "samples") %>% transposer("samples", "otu_id", separated_categories = c("taxon_id", "otu_id"))
```
